The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities

### 2023-12-15 r_r/manuscript_rihe_r_r.Rmd
privhs_df

Earlier in the "enrollment funnel," universities utilize marketing and recruiting interventions to attract nonresident applicants. One widely utilized intervention is off-campus recruiting visits by admissions counselors to local high schools. @RN4758 analyzed off-campus recruiting visits to public high schools by 15 public research universities. Recruiting visits to in-state high schools tended to exhibit modest racial and socioeconomic bias. However, 12 of the 15 universities made more visits to out-of-state high schools than in-state high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. @RN4934 conducted geospatial analyses of out-of-state recruiting visits to Los Angeles and Dallas by four public research universities. She found that universities engage in "recruitment redlining -- the circuitous avoidance of predominantly Black and Latinx communities along recruiting visit paths" (p. X).<!--[CRYSTAL -- FIND PAGE NUMBER] TODO: I don't think I'm able to access the full version of the paper: https://doi.org/10.1080/00221546.2021.2004811--> Although @RN4758 and @RN4934 focused on visits to public high schools, prior research has not considered visits to private schools by public research universities. For example, what is the scale of these visits and what are the characteristics of visited schools?

@@

The data collection sample for the broader project consisted of: all public research-extensive universities as defined by the 2000 Carnegie Classification (N=102), all private universities in the top 100 of U.S. News and World Report National Universities rankings (N=57), and all private colleges in the top 50 of U.S. News and World Report Liberal Arts Colleges rankings (N=47). The rationale for different criteria for inclusion across different university types is because these criteria reflected the interest of a foundation that funded the data collection. This article excludes private liberal arts colleges because having a third institutional type would substantially increase article length. The final analysis sample for this article consists of 15 public research universities and 14 private research universities that we believed posted all varieties of off-campus recruiting events on their admissions website. For these universities, we analyze recruiting visits to private high schools that meet the following criteria: enrolls at least 10 12th grade students, is not primarily a virtual school, and is located in the 50 U.S. states, the District of Columbia, or land regulated by the Bureau of Indian Affairs.

@@

A social network consists of a set of actors -- referred to as "vertices" -- and the connections -- referred to as "network ties" or "edges" between these actors. "One-mode" networks consist of vertices of a "type." For example, the social network of Facebook consists of users (vertices) who are connected to one another via friendship ties (edges). Similarly, the social network of Twitter<!--TODO: Not sure if we should call this X now? haha--> consists of handles (vertices) who are connected to one another by following or being followed (edges).

@@

Figures \@ref(fig:race-pubu-privhs-pubhs) and \@ref(fig:race-privu-privhs-pubhs) show the percentage of 12th graders who identify as Black, Latinx, or Native at visited high schools. Consistent with @RN4504, most private high schools enroll relatively few students from historically underrepresented race/ethnicity groups. The top row in the top panel of Figure \@ref(fig:race-pubu-privhs-pubhs) shows that the sum of Black, Latinx, and Native students comprise less than 20% of enrolled students in over 65%<!--0.6632124--> of schools in our private high school universe. At public high schools (bottom panel), Black/Latinx/Native students comprise less than 20% of enrolled students at around 40%<!--0.4087567--> of all public schools. However, most public research universities and all selective private universities visited private high schools at a higher rate than public high schools<!--TODO: I'm not too sure how to interpret "higher rate than public high schools" - should it be higher rate than proportional rate? Related, is the last sentence supposed to be "disproportionate emphasis on visits to" PRIVATE instead of "public schools"?-->, as seen earlier in Figures \@ref(fig:actual-proportional-pubu) and \@ref(fig:actual-proportional-privu). Even if universities visited schools with racial composition on par with the universe of private high schools, most visited schools would have relatively few Black/Latinx/Native students enrolled. Therefore, a disproportionate emphasis on visits to public schools may shift the racial composition of public universities away from Black/Latinx/Native students.

>>>>

Earlier in the "enrollment funnel," universities utilize marketing and recruiting interventions to attract nonresident applicants. One widely utilized intervention is off-campus recruiting visits by admissions counselors to local high schools. @RN4758 analyzed off-campus recruiting visits to public high schools by 15 public research universities. Recruiting visits to in-state high schools tended to exhibit modest racial and socioeconomic bias. However, 12 of the 15 universities made more visits to out-of-state high schools than in-state high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. @RN4934 conducted geospatial analyses of out-of-state recruiting visits to Los Angeles and Dallas by four public research universities. She found that universities engage in "recruitment redlining -- the circuitous avoidance of predominantly Black and Latinx communities along recruiting visit paths" (p. 585). Although @RN4758 and @RN4934 focused on visits to public high schools, prior research has not considered visits to private schools by public research universities. For example, what is the scale of these visits and what are the characteristics of visited schools?

@@

The data collection sample for the broader project consisted of: all public research-extensive universities as defined by the 2000 Carnegie Classification (N=102), all private universities in the top 100 of the 2020 U.S. News and World Report National Universities rankings (N=57), and all private colleges in the top 50 of the 2020 U.S. News and World Report Liberal Arts Colleges rankings (N=47). The rationale for different criteria for inclusion across different university types is because these criteria reflected the interest of a foundation that funded the data collection. This article excludes private liberal arts colleges because having a third institutional type would substantially increase article length. The final analysis sample for this article consists of 15 public research universities and 14 private research universities that we believed posted all varieties of off-campus recruiting events on their admissions website. For these universities, we analyze recruiting visits to private high schools that meet the following criteria: enrolls at least 10 12th grade students, is not primarily a virtual school, and is located in the 50 U.S. states, the District of Columbia, or land regulated by the Bureau of Indian Affairs.

@@

A social network consists of a set of actors -- referred to as "vertices" -- and the connections -- referred to as "network ties" or "edges" between these actors. "One-mode" networks consist of vertices of a "type." For example, the social network of Facebook consists of users (vertices) who are connected to one another via friendship ties (edges). Similarly, the social network of $X$ (formerly Twitter) consists of handles (vertices) who are connected to one another by following or being followed (edges).

@@

Figures \@ref(fig:race-pubu-privhs-pubhs) and \@ref(fig:race-privu-privhs-pubhs) show the percentage of 12th graders who identify as Black, Latinx, or Native at visited high schools. Consistent with @RN4504, most private high schools enroll relatively few students from historically underrepresented race/ethnicity groups. The top row in the top panel of Figure \@ref(fig:race-pubu-privhs-pubhs) shows that the sum of Black, Latinx, and Native students comprise less than 20% of enrolled students in over 65%<!--0.6632124--> of schools in our private high school universe. At public high schools (bottom panel), Black/Latinx/Native students comprise less than 20% of enrolled students at around 40%<!--0.4087567--> of all public schools. However, most public research universities (Figure \@ref(fig:actual-proportional-pubu)) and all selective private universities (Figure \@ref(fig:actual-proportional-privu)) visited a disproportionate number of private high schools compared to public high schools. Even if universities visited schools with racial composition on par with the universe of private high schools, most visited schools would have relatively few Black/Latinx/Native students enrolled. Therefore, a disproportionate emphasis on visits to public schools may shift the racial composition of public universities away from Black/Latinx/Native students.
